---
title: "Untitled"
author: "Venkata Dhanush Kikkisetti & Druv Jain"
date: "2022-10-31"
output: github_document
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
powerplant<-read_csv('/Users/dhanushkikkisetti/Documents/Data_612/Data 612 project/globalpowerplantdatabasev120/global_power_plant_database.csv')
problems()
powerplant[5504:5510,]
```

# Data cleaning 
```{r}
powerplant<-powerplant%>%
  separate(wepp_id,
           into = c('wepp_id1','wepp_id2','wepp_id3'),
           sep = '|')
powerplant%>%
  summarise(across(everything(),~sum(is.na(.))))
powerplant%>%
  summarise(count(distinct(.)),total=nrow(powerplant))
str(powerplant)
powerplant<-powerplant%>%
  select(-url)
```
# Some basic working on data and creating some summarizations
```{r}
powerplant%>%
 filter(!is.na(generation_gwh_2013)==TRUE)%>%
  ggplot(mapping = aes(x=capacity_mw,y=generation_gwh_2013))+
  geom_point()
  
powerplant%>%
 filter(!is.na(generation_gwh_2014)==TRUE)%>%
  ggplot(mapping = aes(x=capacity_mw,y=generation_gwh_2014))+
  geom_point()
```



# Some basic summary of data
```{r}
unique(powerplant$year_of_capacity_data)
powerplant%>%
  filter(!is.na(year_of_capacity_data))%>%
  group_by(year_of_capacity_data)%>%
  summarise(n=n(),total=mean(capacity_mw))
# Total number of geographical data we have.
powerplant%>%
  filter(!is.na(geolocation_source)==TRUE)%>%
  group_by(geolocation_source)%>%
  summarise(n=n())
# Estimated Power Generation for each country
powerplant%>%
  filter(!is.na(estimated_generation_gwh))%>%
  group_by(country_long)%>%
  summarise(n=n(),average=mean(estimated_generation_gwh))%>%
  arrange(average)
```


# Creating new dataframe that we use for coming code chunk

```{r}
powerplant%>%
  select(country,country_long,primary_fuel,generation_gwh_2013:estimated_generation_gwh)->powerplant_new1

powerplant_new%>%
  summarise(across(everything(),~sum(is.na(.))))
```

# 1) Choosing a particular country and analyzing what type of powerplant has more electricity generation

```{r}
powerplant%>%
  group_by(country,primary_fuel)%>%
  summarise(Total_plants=n(),
            current_generated_2013=sum(generation_gwh_2013,na.rm = TRUE),
            current_generated_2014=sum(generation_gwh_2014,na.rm = TRUE),
            current_generated_2015=sum(generation_gwh_2015,na.rm = TRUE),
            current_generated_2016=sum(generation_gwh_2016,na.rm = TRUE),
            current_generated_2017=sum(generation_gwh_2017,na.rm = TRUE))%>%
            arrange(desc(Total_plants))
options(scipen=3)
# For the Country USA
powerplant_new1%>%
  filter(country=='USA')%>%
  group_by(primary_fuel)%>%
  summarise(total_power=sum(power_generation))%>%
  mutate(primary_fuel=fct_reorder(primary_fuel,total_power))%>%
  ggplot(mapping = aes(x=primary_fuel,y=total_power))+
  geom_col()+
  labs(
    x="Primary_fuel",
    y="Total electricity genenration",
    title="Total power generation for different fossel fuels in USA"
  )+
  theme_bw()

# This is for all Countries
powerplant_new1%>%
  group_by(country_long,primary_fuel)%>%
  summarise(total_power=sum(power_generation))%>%
  mutate(primary_fuel=fct_reorder(primary_fuel,total_power))%>%
  ggplot(mapping = aes(x=primary_fuel,y=log(total_power)))+
  geom_col()+
  labs(
    x="Primary_fuel",
    y="Total electricity genenration",
    title="Total power generation for different fossel fuels"
  )+
  facet_wrap(~country_long)+
  theme_bw()+
  theme(axis.text.x = element_text(angle=90, hjust=1)) 
```
# 2) Analyzing which type of powerplant has more electricity generation for which year and doing a little research on it  

```{r}
#For summarisation table
powerplant%>%
  group_by(primary_fuel)%>%
  summarise(Total_plants=n(),
            current_generated_2013=sum(generation_gwh_2013,na.rm = TRUE),
            current_generated_2014=sum(generation_gwh_2014,na.rm = TRUE),
            current_generated_2015=sum(generation_gwh_2015,na.rm = TRUE),
            current_generated_2016=sum(generation_gwh_2016,na.rm = TRUE),
            current_generated_2017=sum(generation_gwh_2017,na.rm = TRUE))%>%
  arrange(desc(Total_plants))

#For Data Visualization
powerplant_new1%>%
  group_by(primary_fuel,Years)%>%
  filter(primary_fuel!="Storage"&&primary_fuel!="Other")%>%
  summarise(Total_power_generation=sum(power_generation))%>%
  ggplot(mapping = aes(x=Years,y=sqrt(Total_power_generation)))+
  geom_col()+
  facet_wrap(~primary_fuel)+
  theme_bw()+
  labs(
    x="Sqrt transformation of power generation in gigawatts",
    y="Years",
    title="Power geeration for each year with respective to country"
  )
```

# 3) Total number of plants for each fuel
```{r}
powerplant%>%
  group_by(primary_fuel)%>%
  summarise(Total_plants=n())%>%
  mutate(primary_fuel=fct_reorder(primary_fuel,Total_plants))%>%
ggplot(mapping = aes(x=primary_fuel,y=Total_plants))+
  geom_col(fill="orange")+
  theme_bw()+
  labs(
    x="Primary Fuel",
    y="Total power plants",
    title = "Total power plants for each fuel in the world "
  )
```
4) Analyzing if there is a difference in generated electricity and estimated electricity with respect to the country

```{r}
powerplant_new1%>%
  filter(!is.na(estimated_generation_gwh))%>%
  group_by(country_long)%>%
  summarise(Total_power_generation=sum(power_generation),
            Estimated_power_generation=sum(estimated_generation_gwh))
  
```


# 5)Effect of power plant type in generating electricity.  
```{r}
#Visualization for just only one year
powerplant%>%
  group_by(primary_fuel)%>%
  summarise(Total_plants=n(),
            current_generated_2013=sum(generation_gwh_2013,na.rm = TRUE),
            current_generated_2014=sum(generation_gwh_2014,na.rm = TRUE),
            current_generated_2015=sum(generation_gwh_2015,na.rm = TRUE),
            current_generated_2016=sum(generation_gwh_2016,na.rm = TRUE),
            current_generated_2017=sum(generation_gwh_2017,na.rm = TRUE))%>%
  mutate(primary_fuel=fct_reorder(primary_fuel,current_generated_2013))%>%
  ggplot(mapping = aes(x=primary_fuel,y=current_generated_2017))+
  geom_col()+
  theme_bw()

# Visualization for all years

powerplant_new1%>%
  group_by(primary_fuel)%>%
  summarise(Total_power=sum(power_generation))%>%
  mutate(primary_fuel=fct_reorder(primary_fuel,Total_power))%>%
  ggplot(mapping = aes(x=primary_fuel,y=Total_power))+
  geom_col(fill="orange")+
  theme_bw()+
  labs(
    x="Primary Fuel",
    y="Total power generated ",
    title = "Total Electricity generated by each fuel in the world "
  )
```
# 6) Which geolocation_source has higher estimated power generation.

```{r}
powerplant%>%
  filter(!is.na(geolocation_source))%>%
  group_by(geolocation_source)%>%
  summarise(Estimatepower=mean(estimated_generation_gwh,na.rm=TRUE))%>%
  arrange(desc(Estimatepower))%>%
  ggplot(mapping = aes(x=geolocation_source,y=Estimatepower))+
  geom_col()
```
#7) Total power generation for each year with respect to country and top countries with highest power generation
```{r}
powerplant%>%
  select(country,generation_gwh_2013,generation_gwh_2014,generation_gwh_2015,generation_gwh_2016,generation_gwh_2017,estimated_generation_gwh)->powerplant2

powerplant2%>%
  pivot_longer(cols = generation_gwh_2013:generation_gwh_2017,
               names_to = c("Years"),
               values_to = c("power_generation")
               )%>%
  mutate(Years=str_replace_all(Years,"generation_gwh_",' '))%>%
  filter(!is.na(power_generation))->powerplant2

powerplant2%>%
  group_by(country,Years)%>%
  summarise(Total_power_generation=sum(power_generation))->new
ggplot(data=new,mapping = aes(x=Years,y=log(Total_power_generation)))+
  geom_point()+
  facet_wrap(~country)+
    labs(
      x="Total Years",
      y="Total power generation ",
      title = "Total power generation for each year with respective to countriy"
    )

powerplant2%>%
  group_by(country)%>%
  summarise(Total_power_generation=sum(power_generation))%>%
  arrange(desc(Total_power_generation))
```




#8) Estimated vs Actual power generation with respective to primary fuel

```{r}
options(scipen = 10)
powerplant_new1%>%
  filter(!is.na(estimated_generation_gwh))%>%
  ggplot(mapping=aes(x=sqrt(estimated_generation_gwh),y=sqrt(power_generation)))+
  geom_point()+
  labs(
    x="Estimated Generation",
    y="actual power generation"
  )+
  facet_wrap(~primary_fuel)+
  theme_bw()
```

# Using statistical method lets perform statistical tests if current generation increase with increase in years.
## some summary stats for the data set
```{r}
powerplant_new1 %>%
  group_by(Years) %>%
  summarise(Total_power_plants = n(), Average = mean(power_generation), Median = median(power_generation),
            SD = sd(power_generation),  IQR = IQR(power_generation)) %>%
  knitr::kable(digits = 2)
```

## Explantory Data Analysis
# checking for normality
```{r}
ggplot(data=powerplant_new,mapping = aes(x=power_generation))+
  geom_histogram()+
  labs(
    x='power_generation',y='frequency'
  )+
  theme_bw()
```
# power generation from year 2013 to 2017 is increased but average generation is decreased
```{r}
powerplant_new%>%
  ggplot(mapping = aes(x=Years,y=power_generation))+
  geom_col(aes(fill=Years),show.legend = FALSE)+
  labs (title = "Power Generation from year 2013-2017",
        y = "Power generation in gigawatts", x = "Years of generation")
```
# Some Transformations of data which are required for statistical test
```{r}
powerplant_new%>%
  mutate(log_generation=log(power_generation),Years=as.factor(Years))->powerplant_new
powerplant_new%>%
  filter(!is.infinite(log_generation))->powerplant_new

powerplant_new%>%
  summarise(across(everything(),~sum(is.na(.))))
glimpse(powerplant_new)
```

# Histogram of generation of gwh to check if the data is normally distributed
```{r}
ggplot(data=powerplant_new,mapping = aes(x=log_generation))+
  geom_histogram()+
  labs(
    x='logpower_generation',y='frequency'
  )+
  theme_bw()
```

# Some visualisations
```{r}
powerplant_new%>%
  ggplot(mapping = aes(x=Years,y=log_generation))+
  geom_boxplot(aes(fill=Years),show.legend = FALSE)+
  coord_flip()+
  labs (title = "Log Power Generation from year 2013-2017",
        y = "Log Power generation in gigawatts", x = "Years of generation")
```

```{r}
powerplant_new%>%
  group_by(Years)%>%
  summarise(Total_Average=mean(power_generation))%>%
  ggplot(mapping = aes(x=Years,y=Total_Average))+
  geom_col(color='red')+
  theme_bw()+
  ylab("Total Average of current generation")
```

## we can see that data is normally distrubuted and now we can perform anova test considering
Initial Hypothesis as power generation is same for all the years.
Alternative Hypothesis as atleast one year power generation is different 

performing anova by testing equality of mean of power generation by years
```{r}
powerplant_new%>%
  filter(is.nan(log_generation))
anova_model<-aov(log_generation~Years,data=powerplant_new)
anova(anova_model)
```

```{r}
PostHocTest(anova_model, method = "lsd") 
```


